---
title: "Repro"
author: "Julian Flowers"
date: "16 September 2015"
output: html_document
---

This is my R Markdown document for the 1st reproducible research assignment.

##Load and process the data  

1. Load packages  
2. Download and read the data

```{r}
##I'm using lubridate to do some of the date conversions
library(lubridate) 
library(lattice)

## Download, unzip and read in the dataset
## url <-"https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
## download.file(url, destfile = "act.zip")
## unzip("act.zip")

act <- read.csv("activity.csv", header = TRUE, stringsAsFactors = FALSE)
```

See what we have...

```{r}
## Check the dataset
dim(act)
 summary(act)
```

Next, convert the dates to date format 

```{r}
## use the ymd command from lubridate; add an extra column
 act$date1 <- ymd(act$date)
##check its worked
 
 str(act)
```

While we're here, add day of the week using wkday()

```{r}
## use the wkday command from lubridate; add an extra column
act$wkday <- wday(act$date1, label = TRUE)

 summary(act)
```

## Calculate the mean steps per day  

Calculate steps per day and store to a new variable; plot histogram

```{r}
steps_day <- with(act, tapply(steps, date1, sum))
 
     hist(steps_day)

 
 
``` 

###Calculate mean and median steps over all days

```{r}
round(mean(steps_day, na.rm = TRUE),2);median(steps_day, na.rm = TRUE)
```
The mean steps per day was `r round(mean(steps_day, na.rm = TRUE),2)`, and the median was `r median(steps_day, na.rm = TRUE)`




###What is the average daily activity pattern?

```{r}
steps_int <- as.data.frame(cbind(unique(act$interval),with(act, tapply(steps, interval, mean, na.rm = TRUE))))
 colnames(steps_int) <- c("interval", "mean_steps")
 
   with(steps_int, plot(interval, mean_steps, type = 'l'))
``` 

###Which interval contains the max number of steps?

```{r}
which.max(steps_int$mean_steps)
```
  The interval which conatins the maximum number of steps is `r steps_int[104,-2]`.

 
##Impute missing data  
For this I am going to use the mean value for each interval to replace missing data. There are 3 steps  :

* Calculate the mean steps for each interval  
* Replace step NAs for each interval with the mean for that interval
* Combine this as a new field (needs to be unlisted). I have created a new data frame *act2*  



```{r}
act1 <- with(act, tapply(steps, interval, function(x) {ifelse(is.na(x), mean(x, na.rm = TRUE), x)}))
  act2 <- cbind(act, "imp_steps" = unlist(act1))
  str(act2)
```


```{r}
steps1_day <- with(act2, tapply(imp_steps, date1, sum))
  hist(steps1_day)
``` 

```{r}
summary(steps1_day)
``` 

The distribution is now skewed, and the mean is `r signif(mean(steps1_day),5)` and median `r signif(median(steps1_day),5)`. Using this method of imputation, the mean is unchanged but the median has fallen.



###Weekend activity

Finally, create a new variable *weekend* and use the lattice package to plot the data conditioned on weekend

```{r}
  act2$wend <- with(act2, ifelse(wkday == "Sat"| wkday == "Sun", "weekend","weekday"))
```
 
```{r}
 
  xyplot(steps ~ interval|wend,data = act2, lty=1, type = "l")
    
```
